---
title: "Strings_factors"
date: 10/15/19
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
library(p8105.datasets)
```

# string manipulation
```{r}
string_vec=c("my","name", "is", "kelly")

str_detect(string_vec, "m")
```

```{r}
str_replace(string_vec, "m", "M")
```


str_detect means detecting a pattern inside a string 
there are a lot of "str_ " functions, and you can scroll through to see which ones exist in which. 

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think") #starts with i think
str_detect(string_vec, "i think$") #ends with i think
```
starts with "i think" and which ones ends with "i think"
the ^ at the beginning means its the start of the string variable. 

```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Bb]ush") #detect words with capitol B or lowercase b, but not all caps 
```
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]") #a number followed by any letter (uppercase or lowercase letters) 
```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") # the "." represents a character. so like 7:11 should work
```
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[") #if you want to actually use a square bracket or acutally a "." in the function. Here we are just lookng for an open square bracket
```

## pulse data
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = str_replace(visit, "bl", "00m"),
    visit = fct_relevel(visit, str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

print(pulse_data, n = 12)
```
use str_replace function in the mutate step

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  .[[1]] %>%
  html_table() %>%
  slice(-1) %>%
  as_tibble()
```

```{r}
table_marj %>% 
  select(-contains("P value")) %>% 
  pivot_longer(
    -State,
    names_to="age_year",
    values_to="percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep= "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent=as.numeric(percent)
  )
```


## Factor
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex

as.numeric(vec_sex)
## but this gives different results
vec_sex = relevel(vec_sex, ref = "male")
vec_sex

as.numeric(vec_sex)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

if we want waterhole first
```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
# use reorder instead
```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
reorder the name name with tmax. this function is saying look at hte facotrs that we have, and then find which one has the smallest median tmax, and thats one factor. then look at the next smallest median tmax, and so forth 


#airbnb
```{r}
data("nyc_airbnb")

nyc_airbnb %>%
  filter(neighbourhood_group == "Manhattan") %>% 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price, na.rm = TRUE)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```
reorder the neighborhood by price
coord_flip(), takes thigns that were on the X axis now on the Y axis

```{r}
data("nyc_airbnb")

nyc_airbnb %>%
  filter(neighbourhood_group == "Queens") %>% 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price, na.rm = TRUE)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```

what about factors in linear models?
```{r}
weather_df %>% 
  mutate(name=forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhold_WA"))) %>% 
  lm(tmax~name, data=.)
```

```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)

rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))

rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)
```

```{r}
rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Brooklyn"), "HipsterHQ")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() + 
  scale_fill_viridis(discrete = TRUE)

## gives a warning message -- trying to assign a factor variable that assigns a new variable 
# using string replace ALMOST works --> it has now reported it back to string variable instead of keeping it as a vector variable 


```


